Proving linearizability using forward simulations

### 2017-01-25 draftCAV/app-tss.tex
fixed some spaces
TSS appendix revised

%First, we describe the TS-Stack algorithm:
%\begin{lstlisting}
%struct Node{
%  int data;
%  int ts;
%  Node* next;
%  boolean taken;
%};
%Node* pools[maxThreads];
%int TS = 0;   
%
%void push(int x) {
%  Node* n = new Node(x,MAX_INT,
%                        null,false);
%  n->next = pools[myTID];
%  pools[myTID] = n;
%  int i = TS++;
%  n->ts = i;
%}
%int pop() {
% boolean success = false;
% int maxTS = -1;
% Node* youngest = null;
% while ( !success ) {
%   maxTS = -1; youngest = null;
%   for(int i=0; i<maxThreads; i++){
%     Node* n = pools[i];
%     while (n->taken && n->next != n)
%       n = n->next;
%     if(maxTS < n->ts) {
%       maxTS = n->ts; youngest = n;
%     }
%   }
%   if (youngest != null)
%     success=CAS(youngest->taken,
%                       false,true);
% }
% return youngest->data;
%}
%\end{lstlisting}
\begin{figure}[t]
\centering
\includegraphics[width=10cm]{tssPushFlow.pdf}
\includegraphics[width=12cm]{tssPopFlow.pdf}
\vspace{-8mm}
\caption{The flow diagram for the pop and push methods of the Time-Stamped Stack algorithm. The blue points show the control points roughly and the arrows show the possible transitions.}
\label{fig:tssFlow}
\end{figure}

@@

Next, we will show that $\mathit{fs}_2$ is really a Com(pop)-forward simulation relation. Except the trivial base case, we case-split on the transition rules. We first assume $(s, \alpha s') \in \delta_C$ and $t \in \mathit{fs}_2[s]$. Then, we find  corresponding transition $\alpha' \in \Sigma_{AbsS}$ obeying the Com(pop)-forward simulation relation conditions and obtain $t'$ such that $(t, \alpha' t') \in \delta_{AbsS}$  and $t' \in \mathit{fs}_2[s']$.

@@

Let the following derivation rule of $TSS$ be the one for describing $\alpha$:
\begin{mathpar}
    \inferrule{
      \psi
    }{
      s
      \xrightarrow{\alpha} 
      s'
    }
\end{mathpar}
and the following one be the derivation rule of $AbsS$ describing $\alpha'$ if $\alpha' \neq \epsilon$ (equivalently $\alpha' = \alpha$):
\begin{mathpar}
    \inferrule{
      \psi'
    }{
      q
      \xrightarrow{\alpha'} 
      q'
    }
\end{mathpar}
For the cases $\alpha' = \alpha$, we first need to show $\alpha'$ is enabled in state $t$ i.e., $t$ satisfies $\psi'$. If this can not be directly obtained from $s$ satisfies $\psi$ and using one or two obvious conditions on $\mathit{fs}_2$ (since $t \in \mathit{fs}_2[s]$), we show the derivation in the proof. Then, $t'$ is obtained in a unique way since $AbsS$ is deterministic on its alphabet $ \Sigma_{AbsS} = C \cup R \cup Com(pop)$. The, only other thing to show is $t' \in \mathit{fs}_2[t']$. We show this by proving that $t'$ does not violate any of the conditions of the $\mathit{fs}_2$ described above. We only explain why the new instantiations due to the difference between $s'$ and $s$ or the difference between $t'$ and $t$ do not violate the conditions. We skip the instances that we assumed while relating $s$ to $t$.

For the cases in which $\alpha' = \epsilon$, we have $t'=t$ and the only thing to show is $t \in \mathit{fs}_2[s']$. Again, we only explain why the new instantions due to the difference between $s'$ and $s$ do not violate the conditions.
\begin{itemize}
\item[\textsc{init}] $\mathit{fs}_2[{q_0}_{TSS}] =\{{q_0}_{AbsS}\}$
\item[\textsc{call-push}] The same derivation rule of $TSS$ is applied to $t$  to obtain $t'$. The premise of the rule is satisfied by $t$ trivially in the sense explained before. The new vertex $k$ is added to the $O_t$ such that $k$ is maximal, pending and every completed vertex is ordered before $k$ in $t'$. Moreover, $k$ is overlapping with every pending pop. To see that $t' \in \mathit{fs}_2[s']$ we observe the following: \emph{Nodes} condition is preserved because $k \in O_{t'}$. Since the newly added vertex $k$ is maximal and pending in $t'$, \emph{Pend/Comp} condition is preserved. \emph{Frontiers} and \emph{MaximalOV} conditions are not violated since $k$ is added to $ov(p)$ set for every pending pop operation $p$. 
\item[\textsc{push1}] We have $t' = t$ and show $t \in \mathit{fs}_2[s']$.\emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex. \emph{TSOrder} is still preserved. Timestamp of $n_{s'}(k)$ is maximal and every other nodes of push operations with maximal timestamp in $s'$ are pending vertices in $t$. Hence there can be no ordering between those pushes and $k$ in $t$ that can violate \emph{TSOrder}. Moreover, $k$ is maximal in $t$ which means that it cannot be ordered before another push $k'$ of which node has a lower timestamp. \emph{TidOrder} is also satisfied. Since $k$ is ordered after every completed push in $t$ and every other push by the same thread is completed, ordering required by the \emph{TidOrder} is present.
\item[\textsc{push2}] We have $t' = t$ and show $t \in \mathit{fs}_2[s']$. \emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex. One can also see that the \emph{TraverseBefore} condition is preserved. Let the pop $p$ visiting node $m$ and $n_{s'}(k) <^{tr}_{s'} m$. Since $k$ and $p$ are both pending in $s$ and $t \in \mathit{fs}_2[s]$, $k \in ov_t(p)$ (by the \emph{Frontiers} condition). Hence, \emph{TraverseBefore} is preserved. 
\item[\textsc{push3}] We have $t' = t$ and show $t \in \mathit{fs}_2[s']$. We consider two cases: $n_s(k)->taken$ is \texttt{true} or it is \texttt{false}. For the former case, $k \notin O_t$. The only new instantiation we check is $k \notin O_t$ does not violate \emph{Nodes} condition while relating $s'$ to $t$.

@@

\item[\textsc{pop9}]We have $t' = t$ and $t \in \mathit{fs}_2[s']$.
\item[\textsc{ret-pop}] $t'$ is obtained by applying \textsc{ret-pop} rule of $AbsS$ and $t' \in \mathit{fs}_2[s']$.
\end{itemize}
\end{proof}
%\textcolor{red}{TODO: Check ReverseFrontiers for the cases before compop.}
%\textcolor{red}{TODO: Check TraverseBefore for the cases before pop2.}
%\textcolor{red}{TODO: CheckTraverseBeforeNull for the cases befoere pop 5.}
%\textcolor{red}{TODO: Apply Dr. Enea's comments.}
%\textcolor{red}{TODO: Try case split on fs conditions instantiations check.}
%\textcolor{red}{TODO: Add Flow Diagrams to the paper.} 
%\textcolor{red}{TODO: }

>>>>

%First, we describe the TS-Stack algorithm:
%\begin{lstlisting}
%struct Node{
%  int data;
%  int ts;
%  Node* next;
%  boolean taken;
%};
%Node* pools[maxThreads];
%int TS = 0;   
%
%void push(int x) {
%  Node* n = new Node(x,MAX_INT,
%                        null,false);
%  n->next = pools[myTID];
%  pools[myTID] = n;
%  int i = TS++;
%  n->ts = i;
%}
%int pop() {
% boolean success = false;
% int maxTS = -1;
% Node* youngest = null;
% while ( !success ) {
%   maxTS = -1; youngest = null;
%   for(int i=0; i<maxThreads; i++){
%     Node* n = pools[i];
%     while (n->taken && n->next != n)
%       n = n->next;
%     if(maxTS < n->ts) {
%       maxTS = n->ts; youngest = n;
%     }
%   }
%   if (youngest != null)
%     success=CAS(youngest->taken,
%                       false,true);
% }
% return youngest->data;
%}
%\end{lstlisting}
\begin{figure}[t]
\centering
\includegraphics[width=7cm]{tssPushFlow.pdf}
\includegraphics[width=16cm]{tssPopFlow.pdf}
\vspace{-8mm}
\caption{The flow diagram for the pop and push methods of the Time-Stamped Stack algorithm. The blue points show the control points roughly and the arrows show the possible transitions.}
\label{fig:tssFlow}
\end{figure}

@@

Next, we will show that $\mathit{fs}_2$ is really a $C\cup R\cup Com(pop)$-forward simulation relation. Except the trivial base case, we case-split on the transition rules. We first assume $s \xrightarrow{\alpha}_{TSS} s' $ and $t \in \mathit{fs}_2[s]$. Then, we find  corresponding transition $\alpha' \in \Sigma_{AbsS}$ obeying the $C\cup R\cup Com(pop)$-forward simulation relation conditions and obtain $t'$ such that $t \xrightarrow{\alpha'}_{AbsS} st $  and $t' \in \mathit{fs}_2[s']$.

@@

Let the following describe $\alpha$: $\psi \triangleright s\xrightarrow{\alpha}_{TSS} s'$ where $\psi$ is the precondition (guard) that needs to be satisfied for enabling $\alpha$ and $\psi' \triangleright t \xrightarrow{\alpha'}_{AbsS} t'$ describe the $\alpha'$ if $\alpha' \neq \epsilon$ (equivalently $\alpha' = \alpha$).

For the cases $\alpha' = \alpha$, we first need to show $\alpha'$ is enabled in state $t$ i.e., $t$ satisfies $\psi'$. If this can not be directly obtained from the information that $s$ satisfies $\psi$ and using one or two obvious conditions on $\mathit{fs}_2$ (since $t \in \mathit{fs}_2[s]$), we show the derivation in the proof. Then, $t'$ is obtained in a unique way since $AbsS$ is deterministic on its alphabet $ \Sigma_{AbsS} = C \cup R \cup Com(pop)$. The, only other thing to show is $t' \in \mathit{fs}_2[t']$. We show this by proving that $t'$ does not violate any of the conditions of the $\mathit{fs}_2$ described above. Suppose conditions on $\mathit{fs}_2$ are of the form $\forall \overline{k}. guard_{s,t}(\overline{k}) \triangleright \phi_{s,t}(\overline{k})$ where the $overline{k}$ is a vector of operation identifiers and $\phi$ defined on states $s$ and $t$ must hold if the guard defined on $s$ and $t$ holds. We say that a vector $\overline{k_1}$ is a new instantiation of the condtion if $\overline{k_1}$ not satisfy the $guard_{s,t}$ while relating pre-states, but it satisfies $guard_{s',t'}$ while relating post-states and $\phi_{s',t'}$ needs to be checked non-trivially.

We only explain why the new instantiations due to the difference between $s'$ and $s$ or the difference between $t'$ and $t$ do not violate the conditions. We skip the instances that we assumed while relating $s$ to $t$.

For the cases in which $\alpha' = \epsilon$, we have $t'=t$ and the only thing to show is $t \in \mathit{fs}_2[s']$. Again, we only explain why the new instantiations due to the difference between $s'$ and $s$ do not violate the conditions.

In the following, we show that $\mathit{fs}_2$ is a $C \cup R \cup Com(pop)$-forward simulation relation.
\begin{itemize}
\item[\textsc{init}] $\mathit{fs}_2[{q_0}_{TSS}] =\{{q_0}_{AbsS}\}$
\item[\textsc{call-push}] The same derivation rule of $TSS$ is applied to $t$  to obtain $t'$. The premise of the rule is satisfied by $t$ trivially in the sense explained before. The new vertex $k$ is added to the $O_t$ such that $k$ is maximal, pending and every completed vertex is ordered before $k$ in $t'$. Moreover, $k$ is overlapping with every pending pop. To see that $t' \in \mathit{fs}_2[s']$ we observe the following: \emph{Nodes} condition is preserved because $k \in O_{t'}$. Since the newly added vertex $k$ is maximal and pending in $t'$, \emph{Pend/Comp} condition is preserved. \emph{Frontiers} and \emph{MaximalOV} conditions are not violated since $k$ is added to $ov(p)$ set for every pending pop operation $p$. 
\item[\textsc{push1}] We have $t' = t$ and show $t \in \mathit{fs}_2[s']$.\emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex. \emph{TSOrder} is still preserved. Timestamp of $n_{s'}(k)$ is maximal and every other nodes of push operations with maximal timestamp in $s'$ are pending vertices in $t$. Hence there can be no ordering between those pushes and $k$ in $t$ that can violate \emph{TSOrder}. Moreover, $k$ is maximal in $t$ which means that it cannot be ordered before another push $k'$ of which node has a lower timestamp. \emph{TidOrder} is also satisfied. Since $k$ is ordered after every completed push in $t$ and every other push by the same thread is completed, ordering required by the \emph{TidOrder} is present.
\item[\textsc{push2}] We have $t' = t$ and show $t \in \mathit{fs}_2[s']$. \emph{Nodes} and \emph{Pend/Comp} conditions are still satisfied since $k$ remains to be a pending vertex. One can also see that the \emph{TraverseBefore} condition is preserved. Let the pop $p$ visiting node $m$ and $n_{s'}(k) <^{tr}_{s'} m$. Since $k$ and $p$ are both pending in $s$ and $t \in \mathit{fs}_2[s]$, $k \in ov_t(p)$ (by the \emph{Frontiers} condition). Hence, \emph{TraverseBefore} is preserved. 
\item[\textsc{push3}] We have $t' = t$ and show $t \in \mathit{fs}_2[s']$. We consider two cases: $n_s(k)->taken$ is \texttt{true} or it is \texttt{false}. For the former case, $k \notin O_t$. The only new instantiation we check is $k \notin O_t$ does not violate \emph{Nodes} condition while relating $s'$ to $t$.

@@

\item[\textsc{pop9}]We have $t' = t$ and $t \in \mathit{fs}_2[s']$.
\item[\textsc{ret-pop}] $t'$ is obtained by applying \textsc{ret-pop} rule of $AbsS$ and $t' \in \mathit{fs}_2[s']$.
\end{itemize}
\end{proof}
%\textcolor{red}{TODO: Check ReverseFrontiers for the cases before compop.}
%\textcolor{red}{TODO: Check TraverseBefore for the cases before pop2.}
%\textcolor{red}{TODO: CheckTraverseBeforeNull for the cases befoere pop 5.}
%\textcolor{red}{TODO: Apply Dr. Enea's comments.}
%\textcolor{red}{TODO: Try case split on fs conditions instantiations check.}
